Nathan Greenslit Data Analysis Project

### 2023-04-07 code/analysis_code/analysis_README.Rmd
Worked with R2 pearson and wrapped up Correlation plots

This folder contains code for analyzing cleaned data sets and aids in visualization of parameters of interest.\
(1) The dust subfolder contains `daily_dust.qmd`, which looks at dust concentration over the daily time series. It also contains `dust_and_copies.qmd` which combine Vibrio enumeration with dust data.\
(2) The env sub-folder contains `env_par.qmd` that examines temperature and salinity over the time series.\
(3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
(4) The stats folder contains `dust_copies_stats.qmd` that runs models on variables of interest. \

>>>>

This folder contains code for analyzing cleaned data sets and aids in visualization of parameters of interest.\
(1) The dust subfolder contains `daily_dust.qmd`, which looks at dust concentration over the daily time series. It also contains `dust_and_copies.qmd` which combine Vibrio enumeration with dust data.\
(2) The env sub-folder contains `env_par.qmd` that examines temperature and salinity over the time series.\
(3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
(4) The stats folder contains:
- `1_distributiom.qmd`: Checks distribution of variables of interest
- `2_cca.qmd`: Performs cross-correlation analysis to test for lags in time series (between dust input and Vibrio growth)
- `3_correlation.qmd`: Runs Correlation matrices on all variables
- `4_uni_variate.qmd` : runs univariate models on variable and predictors of interest.
- `5_multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)

### 2023-04-07 code/analysis_code/stats/4_uni_variate.qmd
Worked with R2 pearson and wrapped up Correlation plots

```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))

@@

```{r}
fig1<- bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO+1)",
       x = "log(dust)")

fig2<- c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2+2)",
       x = "log(dust)")

fig3<- rd2 %>% ggplot(aes(lag, copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD+2)",
       x = "log(dust)")
```

#### Stack and Save

```{r}
png(file = here("results", "plots", "stats", "lm.png"))
ggarrange(fig1, fig2, fig3 + font("x.text", size = 10), ncol = 1, nrow = 3)
dev.off()
```

@@

>>>>

```{r}
bo2<- bo %>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))

@@

```{r}
fig1<- bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_correlation(method = "pearson")+
  labs(title = "LM DustxCopies (BO+1)",
       x = "log(dust)")

fig2<- c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_correlation(method = "pearson")+
  labs(title = "LM DustxCopies (C2+2)",
       x = "log(dust)")

fig3<- rd2 %>% ggplot(aes(lag, copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_correlation(method = "pearson")+
  labs(title = "LM DustxCopies (RD+2)",
       x = "log(dust)")
```

#### Stac

```{r}
ggarrange(fig1, fig2, fig3 + font("x.text", size = 10), ncol = 1, nrow = 3)
```

@@

<!--# Need to change R2 input here too -->

### 2023-04-07 code/analysis_code/stats/5_multi_variate.qmd
Worked with R2 pearson and wrapped up Correlation plots

### Libraries

@@

### Clean and Trim

@@

### Combine into one big data set

@@

rd_a<- master %>%
  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06')) 
rd_a<- rd_a %>%
mutate(
  lag =Lag(rd_a$t7sum, shift  = 2)) %>% #Make 2 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%
select(!c(date,t7sum,site))
```

>>>>

# Getting Started

#### Libraries

@@

# Cleaning

#### Clean and Trim

@@

#### Make into one data set

@@

rd_a<- master %>%
  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06')) 
rd_a<- rd_a %>%
mutate(
  lag =Lag(rd_a$t7sum, shift  = 1)) %>% #Make 2 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%
select(!c(date,t7sum,site))
```

### 2023-04-07 products/manuscript/Manuscript.qmd
Worked with R2 pearson and wrapped up Correlation plots

    -   Why are my RMSE and AIC so high?

@@

@@

The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\]. Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico. These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each. Dust aerosols can harbor a wide diversity of bacteria \[@kelloggCharacterizationAerosolizedBacteria2004\], fungi \[@ramirez-camejoFungalDiversitySahara2022\], virus-like particles \[@griffinAfricanDesertDust\], and minerals \[@formenti2003\]. These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] . This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].

@@

Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens. Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming. The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V. cholerae*), shellfish-induced gastroenteritis (*V. parahaemolyticus, V. vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V. vulnificus*), and wound and ear infections  (*V. alginolyticus*, *V. vulnificus*) [@baker-austinVibrioSppInfections2018\]. Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V. tubiashii, V. coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V. coralliilyticus*) \[@richardsMortalitiesEasternPacific2015; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003\], and mortality in important aquaculture species such as penaeid shrimp (*V. parahaemolyticus*) and sea breams (*V. harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.

@@

![](images/lm.png){width="329"}

>>>>

    -   Why are my RMSE and AIC so high (see table under *Conclusion* section)?

@@

4.  In my `code/analysis_code/stats/3_uni_variate.qmd` code I was playing around with R2 for my linear models. Previously I have been plotting R2 onto the graphs using `stat_poly_equation()` but then I came across the `stat_correlation()` command. When specifying "pearson" as the method, I get a MUCH higher R2 value for the models (\~0.48 to \~0.70). I have been trying to figure out what `stat_poly_equation` uses as its method and why there is such a difference. At least with dust and copies_mL, I expect Pearson's to be a good method. But is there some serious error that I am making here?

@@

The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year \[@dalmeidaModelSaharanDust1986\]. Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico. These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each. Dust aerosols can harbor a wide diversity of bacteria \[@kelloggCharacterizationAerosolizedBacteria2004\], fungi \[@ramirez-camejoFungalDiversitySahara2022\], virus-like particles \[@griffinAfricanDesertDust\], and minerals \[@formenti2003\]. These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] . This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a\].

@@

Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens. Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming. The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V. cholerae*), shellfish-induced gastroenteritis (*V. parahaemolyticus, V. vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V. vulnificus*), and wound and ear infections  (*V. alginolyticus*, *V. vulnificus*) \[@baker-austinVibrioSppInfections2018\]. Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V. tubiashii, V. coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V. coralliilyticus*) \[@richardsMortalitiesEasternPacific2015; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003\], and mortality in important aquaculture species such as penaeid shrimp (*V. parahaemolyticus*) and sea breams (*V. harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010\] among others.

@@

![](images/Screen%20Shot%202023-04-06%20at%2010.23.28%20PM.png)
